\LARGE \bf Mean Estimation from One-Bit Measurements

### 2021-09-29 proof-lan-distributed.tex
fixed all minor comments from R1
fixed all comments after first revision, sending to John

We begin with an argument on the smoothness properties of the densities,
which is important for our Taylor expansions to come.
\begin{lemma}
  \label{lemma:derivative-bounds}
  Let
  Assumption~\ref{assumption:detection-regions}\eqref{item:lipschitz-density}
%  and \eqref{item:finite-intervals} 
 hold. Then for any $A = \cup_{i = 1}^k
  \{[a_i, b_i]\}$ and $h \in \R$,
  \begin{equation*}
    \left|P_{\theta + h}(A) - P_\theta(A) - \dPtheta(A) h \right| \leq 
    k \cdot \lip(f) h^2,
  \end{equation*}
  where
  \begin{equation}
    \dPtheta(A) = \sum_{i = 1}^k f(a_i - \theta) - f(b_i - \theta).
    \label{eqn:expansion-dPtheta}
  \end{equation}
  Additionally, we have the bounds
  \begin{equation}
    \label{eqn:bound-density-diffs}
    |f(b) - f(a)| \le 2 \sqrt{\lip(f) P([a, b])}
    ~~ \mbox{and} ~~
    |\dPtheta(A)| \le 2 \sqrt{k \lip(f)}.
  \end{equation}
\end{lemma}
\noindent
See Section~\ref{sec:proof-derivative-bounds} for a proof.

@@

For the second, we require a bit more work.
Let $L = \lip(f)$ for shorthand. Let $a < b$.
Then we always have
\begin{equation}
  \label{eqn:interval-length-thing}
  P([a, b])
  \ge \int_a^b f(z) dz
  \ge \int_a^b \max\{f(b) - L (b - z), f(a) - L(z - a), 0\} dz.
\end{equation}
If $f(a) + f(b) \ge L (b - a)$, then the point $\hat{z} =
\frac{a + b}{2} - \frac{f(b) - f(a)}{2L}$ satisfies both $f(b) - L(b -
\hat{z}) \ge 0$ and $f(a) - L(\hat{z} - a) \ge 0$. The
integral~\eqref{eqn:interval-length-thing} then becomes
\begin{align*}
  \lefteqn{\int_a^{\hat{z}}
    (f(a) - L (z - a) dz)
    + \int_{\hat{z}}^b
    (f(b) - L (b - z)) dz} \\
  & = \frac{f(a) + f(b)}{2}
  \left(\frac{b - a}{2}\right)
  - L \left(\frac{b - a}{2} \right)^2
  + \frac{(f(b) - f(a))^2}{4L},
\end{align*}
and using the assumption that $\frac{f(a) + f(b)}{2} \ge L(b - a)$,
we obtain
\begin{align*}
  \frac{(f(b) - f(a))^2}{4L}
  & \le \frac{f(b) + f(b)}{2} \frac{b - a}{2}
  - L \left(\frac{b - a}{2}\right)^2
  + \frac{(f(b) - f(a))^2}{4L}
  \le P([a, b]).
\end{align*}
That is, $|f(b) - f(a)| \le 2 \sqrt{\lip(f) P([a, b])}$.
In the converse case that $f(a) + f(b) \le L(b - a)$, then
the integral~\eqref{eqn:interval-length-thing} becomes
\begin{align*}
  P([a, b])
  & \ge \int_a^{a + \frac{f(a)}{L}}
  (f(a) - L (z - a)) dz
  + \int_{b - \frac{f(b)}{L}}^b (f(b) - L (b - z)) dz \\
  & = \frac{f(a)^2}{L}
  - \frac{f(a)^2}{2L}
  + \frac{f(b)^2}{L}
  - \frac{f(b)^2}{2L},
\end{align*}
so that $\sqrt{f(a)^2 + f(b)^2} \le \sqrt{2 \lip(f) P([a, b])}$.  In sum,
we have demonstrated that always the first
bound~\eqref{eqn:bound-density-diffs} holds.
To show the second inequality in expression~\eqref{eqn:bound-density-diffs},
note that
$\sum_i P([a_i, b_i]) \le 1$, and apply Cauchy-Schwarz.

@@

Our proof follows that of \cite[Thm.~7.2]{VanDerVaart98} closely. We first
demonstrate a type of uniform quadratic mean differentiability
(Definition~\ref{definition:qmd}) for sets $A$ that are finite unions of
intervals. By a Taylor approximation and concavity of
$\sqrt{\cdot}$, we have
\begin{equation*}
  \sqrt{a} + \frac{b}{2 \sqrt{a}} -
  \frac{b^2}{4\sqrt{a}}
  \le \sqrt{a + b} \le \sqrt{a}
  + \frac{b}{2 \sqrt{a}}
\end{equation*}
for any $a > 0$ and $|b| \le 3a/4$. Consequently,
recalling that $\score_\theta(A) = \dPtheta(A) / P_\theta(A)$,
for any $h \in \R$ and $A = \cup_{i = 1}^k \{[t_i^-, t_i^+]\}$ the
union of $k$ intervals, the expansion~\eqref{eqn:expansion-dPtheta} yields
\begin{equation*}
  \left(\sqrt{P_{\theta + h}(A)} -
  \sqrt{P_\theta(A)} - \half h \score_\theta(A) \sqrt{P_\theta(A)}
  \right)^2
  \le
  \left(
  \frac{k \lip(f)}{2 \sqrt{P_\theta(A)}} h^2
  + \frac{(|\dPtheta(A) h| + h^2 \lip(f))^2}{P_\theta(A)^{3/2}}
  \right)^2,
\end{equation*}
valid for $h$ such that
$|\dPtheta(A) h| \le P_\theta(A) / 4$ and
$k h^2 \lip^2(f) \le P_\theta(A) / 4$.
Thus, under
Assumption~\ref{assumption:detection-regions}\eqref{item:finite-intervals},
there exists a numerical constant $C < \infty$ such that
\begin{subequations}
  \label{eqn:h-fourth}
  \begin{align}
    \nonumber \left(\sqrt{P_{\theta + h}(A)} -
    \sqrt{P_\theta(A)} - \half
    h \score_\theta(A) \sqrt{P_\theta(A)}\right)^2
    & \le
    \left(\frac{h^2 k \cdot \lip(f)}{2 \sqrt{P_\theta(A)}}
    + \frac{(|\dPtheta(A) h| + k h^2 \lip(f))^2}{
      P_\theta(A)^{3/2}}\right)^2 \\
    & \le \frac{C}{P_\theta(A)} \left[k^2 \lip^2(f)
      + \score_\theta(A)^2
      + \frac{k^4 h^4 \lip^4(f)}{P_\theta(A)^2}
      \right] \cdot h^4,
  \end{align}
  valid whenever $|\dPtheta(A) h|
  \le P_\theta(A) / 4$ and $k h^2 \lip^2(f) \le P_\theta(A) / 4$,
  and similarly, we have
  \begin{equation}
    \left(\sqrt{P_{\theta + h}(A^c)} -
    \sqrt{P_\theta(A^c)} - \half
    h \score_\theta(A^c) \sqrt{P_\theta(A^c)}\right)^2
    \le \frac{C}{P_\theta(A^c)}
    \left[k^2 \lip^2(f)
      + \score_\theta(A^c)^2
      + \frac{k^4 h^4 \lip^4(f)}{P_\theta(A^c)^2}
      \right] \cdot h^4.
  \end{equation}
\end{subequations}
That is, the family $\{P_\theta\}$ with bit observations $B_n$ satisfies a
uniform type of quadratic-mean differentiability
(Def.~\ref{definition:qmd}).

>>>>

We begin with an argument on the smoothness properties of the densities,
which is important for our Taylor expansions to come.
\begin{lemma}
  \label{lemma:derivative-bounds}
  Let Assumption~\ref{assumption:detection-regions}\eqref{item:lipschitz-density} hold. Then for any $A = \cup_{i = 1}^k
  [a_i, b_i]$ and $h \in \R$,
  \begin{align}
    \left|P_{\theta + h}(A) - P_\theta(A) - \dPtheta(A) h \right| \leq 
    k \cdot \lip(f) h^2,
    \label{eqn:expansion-dPtheta}
  \end{align}
  where
  \begin{align*}
    \dPtheta(A) = \sum_{i = 1}^k f(a_i - \theta) - f(b_i - \theta).
  \end{align*}
  Additionally, we have the bounds
  \begin{align}
    \label{eqn:bound-density-diffs}
    |f(b) - f(a)| \le 2 \sqrt{\lip(f) P([a, b])}
    ~~ \mbox{and} ~~
    |\dPtheta(A)| \le 2 \sqrt{k \lip(f)}.
  \end{align}
\end{lemma}
\noindent
See Section~\ref{sec:proof-derivative-bounds} for a proof.

@@

For the second, we require a bit more work.
Let $L = \lip(f)$ for shorthand. Let $a < b$.
Then we always have
\begin{equation}
  \label{eqn:interval-length-thing}
  P([a, b])
  \ge \int_a^b f(z) dz
  \ge \int_a^b \max\{f(b) - L (b - z), f(a) - L(z - a), 0\} dz.
\end{equation}
If $f(a) + f(b) \ge L (b - a)$, then the point $\hat{z} =
\frac{a + b}{2} - \frac{f(b) - f(a)}{2L}$ satisfies both $f(b) - L(b -
\hat{z}) \ge 0$ and $f(a) - L(\hat{z} - a) \ge 0$. The
integral~\eqref{eqn:interval-length-thing} then becomes
\begin{align*}
  \lefteqn{\int_a^{\hat{z}}
    (f(a) - L (z - a) dz)
    + \int_{\hat{z}}^b
    (f(b) - L (b - z)) dz} \\
  & = \frac{f(a) + f(b)}{2}
  \left(\frac{b - a}{2}\right)
  - L \left(\frac{b - a}{2} \right)^2
  + \frac{(f(b) - f(a))^2}{4L},
\end{align*}
and using the assumption that $\frac{f(a) + f(b)}{2} \ge L(b - a)$,
we obtain
\begin{align*}
  \frac{(f(b) - f(a))^2}{4L}
  & \le \frac{f(b) + f(b)}{2} \frac{b - a}{2}
  - L \left(\frac{b - a}{2}\right)^2
  + \frac{(f(b) - f(a))^2}{4L}
  \le P([a, b]).
\end{align*}
That is, $|f(b) - f(a)| \le 2 \sqrt{\lip(f) P([a, b])}$.
In the converse case that $f(a) + f(b) \le L(b - a)$, then
the integral~\eqref{eqn:interval-length-thing} becomes
\begin{align*}
  P([a, b])
  & \ge \int_a^{a + \frac{f(a)}{L}}
  (f(a) - L (z - a)) dz
  + \int_{b - \frac{f(b)}{L}}^b (f(b) - L (b - z)) dz \\
  & = \frac{f(a)^2}{L}
  - \frac{f(a)^2}{2L}
  + \frac{f(b)^2}{L}
  - \frac{f(b)^2}{2L},
\end{align*}
so that 
\begin{align*}
\newtext{\frac{f(a) + f(b)}{\sqrt{2}} \le} \sqrt{f(a)^2 + f(b)^2} \le \sqrt{2 \lip(f) P([a, b])},
\end{align*}
\newtext{where the left inequality follows from concavity of $\sqrt{\cdot}$.} In sum, we have demonstrated that always the first
bound~\eqref{eqn:bound-density-diffs} holds.
To show the second inequality in expression~\eqref{eqn:bound-density-diffs},
note that
$\sum_i P([a_i, b_i]) \le 1$, and apply Cauchy-Schwarz.

@@

Our proof follows that of \cite[Thm.~7.2]{VanDerVaart98} closely. We first
demonstrate a type of uniform quadratic mean differentiability
(Definition~\ref{definition:qmd}) for sets $A$ that are finite unions of
intervals. By a Taylor approximation and concavity of
$\sqrt{\cdot}$, we have
\begin{equation*}
  \sqrt{a} + \frac{b}{2 \sqrt{a}} -
  \frac{b^2}{4a^{\newtext{3/2}}}
  \le \sqrt{a + b} \le \sqrt{a}
  + \frac{b}{2 \sqrt{a}}
\end{equation*}
for any $a > 0$ and $|b| \le 3a/4$. Consequently,
recalling that $\score_\theta(A) = \dPtheta(A) / P_\theta(A)$,
for any $h \in \R$ and $A = \cup_{i = 1}^k [t_i^-, t_i^+]$ the
union of $k$ intervals, the expansion~\eqref{eqn:expansion-dPtheta} yields
\begin{equation*}
  \left(\sqrt{P_{\theta + h}(A)} -
  \sqrt{P_\theta(A)} - \half h \score_\theta(A) \sqrt{P_\theta(A)}
  \right)^2
  \le
  \left(
  \frac{k \lip(f)}{2 \sqrt{P_\theta(A)}} h^2
  + \frac{(|\dPtheta(A) h| + h^2 \lip(f))^2}{P_\theta(A)^{3/2}}
  \right)^2,
\end{equation*}
valid for $h$ such that
$|\dPtheta(A) h| \le P_\theta(A) / 4$ and
$k h^2 \lip^2(f) \le P_\theta(A) / 4$.
Thus, under
Assumption~\ref{assumption:detection-regions}\eqref{item:finite-intervals},
there exists a numerical constant $C < \infty$ such that
\begin{subequations}
  \label{eqn:h-fourth}
  \begin{align}
    \nonumber \left(\sqrt{P_{\theta + h}(A)} -
    \sqrt{P_\theta(A)} - \half
    h \score_\theta(A) \sqrt{P_\theta(A)}\right)^2
    & \le
    \left(\frac{h^2 k \cdot \lip(f)}{2 \sqrt{P_\theta(A)}}
    + \frac{(|\dPtheta(A) h| + k h^2 \lip(f))^2}{
      P_\theta(A)^{3/2}}\right)^2 \\
    & \le \frac{C}{P_\theta(A)} \left[k^2 \lip^2(f)
      + \score_\theta(A)^2
      + \frac{k^4 h^4 \lip^4(f)}{P_\theta(A)^2}
      \right] \cdot h^4,
  \end{align}
  valid whenever $|\dPtheta(A) h|
  \le P_\theta(A) / 4$ and $k h^2 \lip^2(f) \le P_\theta(A) / 4$,
  and similarly, we have
  \begin{equation}
    \left(\sqrt{P_{\theta + h}(A^c)} -
    \sqrt{P_\theta(A^c)} - \half
    h \score_\theta(A^c) \sqrt{P_\theta(A^c)}\right)^2
    \le \frac{C}{P_\theta(A^c)}
    \left[k^2 \lip^2(f)
      + \score_\theta(A^c)^2
      + \frac{k^4 h^4 \lip^4(f)}{P_\theta(A^c)^2}
      \right] \cdot h^4.
  \end{equation}
\end{subequations}
That is, the family $\{P_\theta\}$ with bit observations $B_n$ satisfies a
uniform type of quadratic-mean differentiability
(Def.~\ref{definition:qmd}).
